---
title: "Models cross-sectional dimensions"
output: html_document
date: "2026-03-03"
---

```{r libraries}
library(tidyverse)
library(brms)
library(loo)
library(ggplot2)
library(patchwork)
```

```{r dataset}
d_scales <- read.csv("data/d_scales.csv")
```

```{r}
# Bayesian McKelvey-Zavoina R2 ------------------------------------------------
# Bayesian version of McKelvey and Zavoina's pseudo-R2 for binary and ordinal
# brms models (McKelvey and Zavoina, 1975). See also Gelman et al. (2018). This
# pseudo-R2 closely approximates the R2 that would have been obtained if a
# linear model had have been run on observations of the continuous latent
# variable underlying the discrete responses (Veall and Zimmermann, 1992; Hagle
# and Mitchell, 1992; Veall and Zimmermann, 1994).

#from here: https://discourse.mc-stan.org/t/r-2-calculation-for-brm-model-with-cumulative-family-type/14752/3, retrieved on 28th May 2025

Bayes_R2_MZ <- function(fit, ...) {
  y_pred <- fitted(fit, scale = "linear", summary = FALSE, ...)
  var_fit <- apply(y_pred, 1, var)
  if (fit$formula$family$family == "cumulative" ||
      fit$formula$family$family == "bernoulli") {
    if (fit$formula$family$link == "probit" || 
        fit$formula$family$link == "probit_approx") {
      var_res <- 1
    }
    else if (fit$formula$family$link == "logit") {
      var_res <- pi^2 / 3 
    }
  } 
  else {
    sum_fit <- summary(fit)
    sig_res <- sum_fit$spec_pars["sigma", "Estimate"]
    var_res <- sig_res^2
  } 
  R2_MZ <- var_fit / (var_fit + var_res)
  print(
    data.frame(
      Estimate = mean(R2_MZ), 
      Est.Error = sd(R2_MZ), 
      "l-95% CI" = quantile(R2_MZ, 0.025),
      "u-95% CI" = quantile(R2_MZ, 0.975),
      row.names = "Bayes_R2_MZ", 
      check.names = FALSE), 
    digits = 2)
}
```

# Coop 4

```{r}
d_scales %>% 
  ggplot(aes(x=coop_5_2))+
  geom_histogram(fill="pink", position=position_nudge(x=-0.1), binwidth=0.07)+
  geom_histogram(data=d_scales, aes(x=coop_5_3), fill="aquamarine", binwidth=0.07)+
  geom_histogram(data=d_scales, aes(x=coop_5_4), position=position_nudge(x=0.1), fill="darkred", binwidth=0.07)
```

```{r}
d_long_c4 <- d_scales %>% 
  pivot_longer(cols=c("coop_5_2","coop_5_3","coop_5_4")) %>% 
  rename(item_ca=name,
         ca=value)
```

## Compare model with vs without different thresholds

```{r}
prior_empty <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6),
              #random intercepts => so they can vary across 0
              prior(exponential(1), class = sd))

prior_empty_T <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1, group=coop_5_2),
              prior(normal(-0.57,2), class=Intercept, coef=2, group=coop_5_2),
              prior(normal(-0.18,2), class=Intercept, coef=3, group=coop_5_2),
              prior(normal(0.18,2), class=Intercept, coef=4, group=coop_5_2),
              prior(normal(0.57,2), class=Intercept, coef=5, group=coop_5_2),
              prior(normal(1.07,2), class=Intercept, coef=6, group=coop_5_2),
              #second item
              prior(normal(-1.07,2), class=Intercept, coef=1, group=coop_5_3),
              prior(normal(-0.57,2), class=Intercept, coef=2, group=coop_5_3),
              prior(normal(-0.18,2), class=Intercept, coef=3, group=coop_5_3),
              prior(normal(0.18,2), class=Intercept, coef=4, group=coop_5_3),
              prior(normal(0.57,2), class=Intercept, coef=5, group=coop_5_3),
              prior(normal(1.07,2), class=Intercept, coef=6, group=coop_5_3),
              #third item
              prior(normal(-1.07,2), class=Intercept, coef=1, group=coop_5_4),
              prior(normal(-0.57,2), class=Intercept, coef=2, group=coop_5_4),
              prior(normal(-0.18,2), class=Intercept, coef=3, group=coop_5_4),
              prior(normal(0.18,2), class=Intercept, coef=4, group=coop_5_4),
              prior(normal(0.57,2), class=Intercept, coef=5, group=coop_5_4),
              prior(normal(1.07,2), class=Intercept, coef=6, group=coop_5_4),
              #random intercepts => so they can vary across 0
              prior(exponential(1), class = sd))

#prior predictive check
c4_ppc <- brm(data=d_long_c4,
      family=cumulative(),
      prior=prior_empty,
      ca ~ (1|ID)+(1|item_ca),
      iter=1500, warmup=500, chains=4,
      sample_prior = "only",
      seed=1)

c4_ppc_T <- brm(data=d_long_c4,
      family=cumulative(),
      prior=prior_empty_T,
      ca | thres(gr = item_ca) ~ (1|ID)+(1|item_ca),
      iter=1500, warmup=500, chains=4,
      sample_prior = "only",
      seed=1)

c4_ppc_probit <- brm(data=d_long_c4,
      family=cumulative(link=probit),
      prior=prior_empty,
      ca ~ (1|ID)+(1|item_ca),
      iter=1500, warmup=500, chains=4,
      sample_prior = "only",
      seed=1)

c4_ppc_probit_T <- brm(data=d_long_c4,
      family=cumulative(link=probit),
      prior=prior_empty_T,
      ca | thres(gr = item_ca) ~ (1|ID)+(1|item_ca),
      iter=1500, warmup=500, chains=4,
      sample_prior = "only",
      seed=1)

pp_check(c4_ppc, type="bars")
pp_check(c4_ppc_probit, type="bars")
pp_check(c4_ppc_T, type="bars")
pp_check(c4_ppc_probit_T, type="bars") #this may seem a bit more accurate

#model comparison
c4_empty <- brm(data=d_long_c4,
      family=cumulative(),
      prior=prior_empty,
      ca ~ (1|ID)+(1|item_ca),
      iter=4000, warmup=1000, chains=4, cores=3, threads=threading(10),
      seed=1)

c4_empty_probit <- brm(data=d_long_c4,
      family=cumulative(link=probit),
      prior=prior_empty,
      ca ~ (1|ID)+(1|item_ca),
      iter=4000, warmup=1000, chains=4, cores=3, threads=threading(10),
      seed=1)

c4_empty_T <- brm(data=d_long_c4,
      family=cumulative(),
      prior=prior_empty_T,
      ca | thres(gr = item_ca) ~ (1|ID)+(1|item_ca),
      iter=4000, warmup=1000, chains=4, cores=3, threads=threading(10),
      seed=1)

c4_empty_probit_T <- brm(data=d_long_c4,
      family=cumulative(),
      prior=prior_empty_T,
      ca | thres(gr = item_ca) ~ (1|ID)+(1|item_ca),
      iter=4000, warmup=1000, chains=4, cores=3, threads=threading(10),
      seed=1)

saveRDS(c4_empty, "fits/c4_empty.rds")
saveRDS(c4_empty_T, "fits/c4_empty_T.rds")

pp_check(c4_empty, type="bars")
pp_check(c4_empty_T, type="bars")
loo(c4_empty, c4_empty_T)

Bayes_R2_MZ(c4_empty)
Bayes_R2_MZ(c4_empty_T)
```

Since the two models are essentially the same, there is no need to estimate a different threshold for every item

## M0

```{r}
prior_m0 <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6),
              #random intercepts => so they can vary across 0
              prior(exponential(1), class=sd),
              #beta coefficients = no expectations since sd variables
              prior(normal(0,0.5), class=b))

c4_m0 <- brm(data=d_long_c4,
      family=cumulative(),
      prior=prior_m0,
      ca ~ gender_m_w + age_s + mo(pol_orientation) + mo(education)+ mo(n_children) + income_s + (1|house)+(1|ID)+(1|item_ca),
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)

saveRDS(c4_m0, "fits/c4_m0.rds")

pp_check(c4_m0, type="bars")
pp_check(c4_m0,type = "hist", ndraws = 11, binwidth = 1) +
  scale_x_continuous(breaks = 1:7)

summary(c4_m0)

loo(c4_m0) #no high Pk
loo(c4_m0,c4_empty)
Bayes_R2_MZ(c4_m0)
```

## M1

```{r}
prior_m1 <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6),
              #random intercepts => so they can vary across 0
              prior(normal(0, 1), class = sd),
              #beta coefficients = no expectations since sd variables
              prior(normal(0,0.5), class=b, coef=gender_m_w),
              prior(normal(0,0.5), class=b, coef=age_s),
              prior(normal(0,0.5), class=b, coef=income_s),
              prior(normal(0,0.5), class=b, coef=mon_children),
              prior(normal(0,0.5), class=b, coef=mopol_orientation),
              prior(normal(0,0.5), class=b, coef=moeducation),
              #predictors
              prior(normal(0.37,0.2), class=b, coef=si_h_s),
              prior(normal(0.33,0.19), class=b, coef=mor_h_s),
              prior(normal(0.35,0.18), class=b, coef=ce_h_s),
              prior(normal(0.35,0.2), class=b, coef=an_h_s))

c4_m1 <- brm(data=d_long_c4,
      family=cumulative(),
      prior=prior_m1,
      ca ~ gender_m_w + age_s + mo(pol_orientation) + mo(education)+ mo(n_children) + income_s + 
        si_h_s + mor_h_s + ce_h_s + an_h_s +
        (1|house)+(1|ID)+(1|item_ca),
      iter=4000, warmup=1000, chains=4,
      control=list(adapt_delta=0.90),
      seed=1) #to check for the dataset: re-run this with the longer dataset and check the posterior width

saveRDS(c4_m1, "fits/c4_m1.rds")

summary(c4_m1)
loo(c4_m0,c4_m1)
Bayes_R2_MZ(c4_m1)
```

## M2

Adding the expected mediation paths

```{r}
d_long_c4_m2 <- d_long_c4 %>% 
  pivot_longer(cols=starts_with("ce_house")) %>% 
  rename(item_ce=name,
         ce=value) %>% 
  pivot_longer(cols=starts_with("anger_house")) %>% 
  rename(item_an=name,
         an=value)

main_model <- bf(ca ~ gender_m_w + age_s + mo(pol_orientation) + mo(education)+ mo(n_children) + income_s + 
                   si_h_s + mor_h_s + ce_h_s + an_h_s +
                   (1|house)+(1|ID)+item_ca)

mediation_an <- bf(an ~ si_h_s + mor_h_s + (1|ID) + item_an)
mediation_ce <- bf(ce ~ si_h_s + mor_h_s + (1|ID) + item_ce)

prior_m2 <- c(
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=ca),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=ca),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=ca),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=ca),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=ca),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=ca),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=an),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=an),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=an),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=an),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=an),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=an),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=ce),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=ce),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=ce),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=ce),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=ce),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=ce),
  #random intercepts => so they can vary across 0
  prior(normal(0, 1), class=sd, resp=ca),
  prior(normal(0, 1), class=sd, resp=an),
  prior(normal(0, 1), class=sd, resp=ce),
  #beta coefficients = no expectations since sd variables
  prior(normal(0,0.5), class=b, coef=gender_m_w, resp=ca),
  prior(normal(0,0.5), class=b, coef=age_s, resp=ca),
  prior(normal(0,0.5), class=b, coef=income_s, resp=ca),
  prior(normal(0,0.5), class=b, coef=mon_children, resp=ca),
  prior(normal(0,0.5), class=b, coef=mopol_orientation, resp=ca),
  prior(normal(0,0.5), class=b, coef=moeducation, resp=ca),
  #predictors
  prior(normal(0.37,0.2), class=b, coef=si_h_s, resp=ca),
  prior(normal(0.33,0.19), class=b, coef=mor_h_s, resp=ca),
  prior(normal(0.35,0.18), class=b, coef=ce_h_s, resp=ca),
  prior(normal(0.35,0.2), class=b, coef=an_h_s, resp=ca),
  #for the mediation models
  prior(normal(0,0.5), class=b, resp=an),
  prior(normal(0,0.5), class=b, resp=ce))

c4_m2 <- brm(data=d_long_c4_m2,
             family=cumulative(),
             main_model + mediation_an + mediation_ce,
             iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
             prior=prior_m2,
             seed=1)

summary(c4_m2)
```

See if quicker

```{r}
main_model <- bf(ca ~ 
                   si_h_s + mor_h_s + ce_h_s + an_h_s +
                   house+(1|ID)+item_ca)

mediation_an <- bf(an ~ si_h_s + mor_h_s + (1|ID) + item_an)
mediation_ce <- bf(ce ~ si_h_s + mor_h_s + (1|ID) + item_ce)

prior_m2 <- c(
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=ca),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=ca),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=ca),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=ca),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=ca),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=ca),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=an),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=an),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=an),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=an),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=an),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=an),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=ce),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=ce),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=ce),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=ce),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=ce),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=ce),
  #random intercepts => so they can vary across 0
  prior(normal(0, 1), class=sd, resp=ca),
  prior(normal(0, 1), class=sd, resp=an),
  prior(normal(0, 1), class=sd, resp=ce),
  #predictors
  prior(normal(0.37,0.2), class=b, coef=si_h_s, resp=ca),
  prior(normal(0.33,0.19), class=b, coef=mor_h_s, resp=ca),
  prior(normal(0.35,0.18), class=b, coef=ce_h_s, resp=ca),
  prior(normal(0.35,0.2), class=b, coef=an_h_s, resp=ca),
  #for the mediation models
  prior(normal(0,0.5), class=b, resp=an),
  prior(normal(0,0.5), class=b, resp=ce))

c4_m2 <- brm(data=d_long_c4_m2,
             family=cumulative(),
             main_model + mediation_an + mediation_ce,
             iter=500, warmup=40, chains=2,
             prior=prior_m2,
             seed=1)

summary(c4_m2)
c4_m2_test <- readRDS("fits/c4_m2_test_virtual_workspace.rds")
summary(c4_m2_test)
```

## Model comparison

```{r}
c4_empty <- readRDS("fits/c4_empty.rds")
c4_m0 <- readRDS("fits/c4_m0.rds")
c4_m1 <- readRDS("fits/c4_m1.rds")

(loo(c4_empty,c4_m0,c4_m1))$diffs %>% 
  as.data.frame() %>% 
  mutate(ratio=elpd_diff/se_diff)

summary(c4_empty)
summary(c4_m0)
summary(c4_m1)
```

# Coop 5

```{r}
d_scales %>% 
  ggplot(aes(x=coop_6_1))+
  geom_histogram(fill="pink", position=position_nudge(x=-0.15), binwidth=0.07)+
  geom_histogram(data=d_scales, aes(x=coop_6_2), fill="aquamarine", binwidth=0.07, position=position_nudge(x=-0.05))+
  geom_histogram(data=d_scales, aes(x=coop_6_3), fill="lightyellow", binwidth=0.07, position=position_nudge(x=0.05))+
  geom_histogram(data=d_scales, aes(x=coop_6_5), position=position_nudge(x=0.15), fill="darkorange", binwidth=0.07)
```

```{r}
d_long_c5 <- d_scales %>% 
  pivot_longer(cols=c("coop_6_1","coop_6_2","coop_6_3","coop_6_5")) %>% 
  rename(item_ca=name,
         ca=value)
```

## Compare model with vs without different thresholds

```{r}
prior_empty <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6),
              #random intercepts => so they can vary across 0
              prior(normal(0, 1), class = sd))

prior_empty_T <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1, group=coop_6_1),
              prior(normal(-0.57,2), class=Intercept, coef=2, group=coop_6_1),
              prior(normal(-0.18,2), class=Intercept, coef=3, group=coop_6_1),
              prior(normal(0.18,2), class=Intercept, coef=4, group=coop_6_1),
              prior(normal(0.57,2), class=Intercept, coef=5, group=coop_6_1),
              prior(normal(1.07,2), class=Intercept, coef=6, group=coop_6_1),
              #second item
              prior(normal(-1.07,2), class=Intercept, coef=1, group=coop_6_2),
              prior(normal(-0.57,2), class=Intercept, coef=2, group=coop_6_2),
              prior(normal(-0.18,2), class=Intercept, coef=3, group=coop_6_2),
              prior(normal(0.18,2), class=Intercept, coef=4, group=coop_6_2),
              prior(normal(0.57,2), class=Intercept, coef=5, group=coop_6_2),
              prior(normal(1.07,2), class=Intercept, coef=6, group=coop_6_2),
              #third item
              prior(normal(-1.07,2), class=Intercept, coef=1, group=coop_6_3),
              prior(normal(-0.57,2), class=Intercept, coef=2, group=coop_6_3),
              prior(normal(-0.18,2), class=Intercept, coef=3, group=coop_6_3),
              prior(normal(0.18,2), class=Intercept, coef=4, group=coop_6_3),
              prior(normal(0.57,2), class=Intercept, coef=5, group=coop_6_3),
              prior(normal(1.07,2), class=Intercept, coef=6, group=coop_6_3),
              #fourth item
              prior(normal(-1.07,2), class=Intercept, coef=1, group=coop_6_5),
              prior(normal(-0.57,2), class=Intercept, coef=2, group=coop_6_5),
              prior(normal(-0.18,2), class=Intercept, coef=3, group=coop_6_5),
              prior(normal(0.18,2), class=Intercept, coef=4, group=coop_6_5),
              prior(normal(0.57,2), class=Intercept, coef=5, group=coop_6_5),
              prior(normal(1.07,2), class=Intercept, coef=6, group=coop_6_5),
              #random intercepts => so they can vary across 0
              prior(normal(0, 1), class = sd))

#prior predictive check
c5_ppc <- brm(data=d_long_c5,
      family=cumulative(),
      prior=prior_empty,
      ca ~ (1|ID)+(1|item_ca),
      iter=1500, warmup=500, chains=4,
      sample_prior = "only",
      seed=1)

c5_ppc_T <- brm(data=d_long_c5,
      family=cumulative(),
      prior=prior_empty_T,
      ca | thres(gr = item_ca) ~ (1|ID)+(1|item_ca),
      iter=1500, warmup=500, chains=4,
      sample_prior = "only",
      seed=1)

pp_check(c5_ppc, type="bars")
pp_check(c5_ppc_T, type="bars") #this may seem a bit more accurate

#model comparison
c5_empty <- brm(data=d_long_c5,
      family=cumulative(),
      prior=prior_empty,
      ca ~ (1|ID)+(1|item_ca),
      iter=1500, warmup=500, chains=4,
      seed=1)

c5_empty_T <- brm(data=d_long_c5,
      family=cumulative(),
      prior=prior_empty_T,
      ca | thres(gr = item_ca) ~ (1|ID)+(1|item_ca),
      iter=1500, warmup=500, chains=4,
      seed=1)

saveRDS(c5_empty, "fits/c5_empty.rds")
saveRDS(c5_empty_T, "fits/c5_empty_T.rds")

pp_check(c5_empty, type="bars")
pp_check(c5_empty_T, type="bars")

(loo(c5_empty, c5_empty_T))$diffs %>% 
  as.data.frame() %>% 
  mutate(ratio=elpd_diff/se_diff)

Bayes_R2_MZ(c5_empty)
Bayes_R2_MZ(c5_empty_T)
```

Since the two models are essentially the same, there is no need to estimate a different threshold for every item

## M0

```{r}
prior_m0 <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6),
              #random intercepts => so they can vary across 0
              prior(normal(0, 1), class=sd),
              #beta coefficients = no expectations since sd variables
              prior(normal(0,0.5), class=b))

c5_m0 <- brm(data=d_long_c5,
      family=cumulative(),
      prior=prior_m0,
      ca ~ gender_m_w + age_s + mo(pol_orientation) + mo(education)+ mo(n_children) + income_s + (1|house)+(1|ID)+(1|item_ca),
      iter=4000, warmup=1000, chains=4,
      seed=1)

saveRDS(c5_m0, "fits/c5_m0.rds")

pp_check(c5_m0, type="bars")
pp_check(c5_m0,type = "hist", ndraws = 11, binwidth = 1) +
  scale_x_continuous(breaks = 1:7)

summary(c5_m0)

loo(c5_m0) #no high Pk
loo(c5_m0,c5_empty)
Bayes_R2_MZ(c5_m0)
```

## M1

```{r}
prior_m1 <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6),
              #random intercepts => so they can vary across 0
              prior(normal(0, 1), class = sd),
              #beta coefficients = no expectations since sd variables
              prior(normal(0,0.5), class=b, coef=gender_m_w),
              prior(normal(0,0.5), class=b, coef=age_s),
              prior(normal(0,0.5), class=b, coef=income_s),
              prior(normal(0,0.5), class=b, coef=mon_children),
              prior(normal(0,0.5), class=b, coef=mopol_orientation),
              prior(normal(0,0.5), class=b, coef=moeducation),
              #predictors
              prior(normal(0.37,0.2), class=b, coef=si_h_s),
              prior(normal(0.33,0.19), class=b, coef=mor_h_s),
              prior(normal(0.35,0.18), class=b, coef=ce_h_s),
              prior(normal(0.35,0.2), class=b, coef=an_h_s))

c5_m1 <- brm(data=d_long_c5,
      family=cumulative(),
      prior=prior_m1,
      ca ~ gender_m_w + age_s + mo(pol_orientation) + mo(education)+ mo(n_children) + income_s + 
        si_h_s + mor_h_s + ce_h_s + an_h_s +
        (1|house)+(1|ID)+(1|item_ca),
      iter=4000, warmup=1000, chains=4,
      seed=1)

saveRDS(c5_m1, "fits/c5_m1.rds")

pp_check(c5_m1, type="bars")

summary(c5_m1)
loo(c5_m0,c5_m1)
Bayes_R2_MZ(c5_m1)
```

## M2

Adding the expected mediation paths

```{r}
d_long_c4_m2 <- d_long_c4 %>% 
  pivot_longer(cols=starts_with("ce_house")) %>% 
  rename(item_ce=name,
         ce=value) %>% 
  pivot_longer(cols=starts_with("anger_house")) %>% 
  rename(item_an=name,
         an=value)

main_model <- bf(ca ~ gender_m_w + age_s + mo(pol_orientation) + mo(education)+ mo(n_children) + income_s + 
                   si_h_s + mor_h_s + ce_h_s + an_h_s +
                   (1|house)+(1|ID)+item_ca)

mediation_an <- bf(an ~ si_h_s + mor_h_s + (1|ID) + item_an)
mediation_ce <- bf(ce ~ si_h_s + mor_h_s + (1|ID) + item_ce)

prior_m2 <- c(
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=ca),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=ca),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=ca),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=ca),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=ca),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=ca),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=an),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=an),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=an),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=an),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=an),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=an),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=ce),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=ce),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=ce),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=ce),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=ce),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=ce),
  #random intercepts => so they can vary across 0
  prior(normal(0, 1), class=sd, resp=ca),
  prior(normal(0, 1), class=sd, resp=an),
  prior(normal(0, 1), class=sd, resp=ce),
  #beta coefficients = no expectations since sd variables
  prior(normal(0,0.5), class=b, coef=gender_m_w, resp=ca),
  prior(normal(0,0.5), class=b, coef=age_s, resp=ca),
  prior(normal(0,0.5), class=b, coef=income_s, resp=ca),
  prior(normal(0,0.5), class=b, coef=mon_children, resp=ca),
  prior(normal(0,0.5), class=b, coef=mopol_orientation, resp=ca),
  prior(normal(0,0.5), class=b, coef=moeducation, resp=ca),
  #predictors
  prior(normal(0.37,0.2), class=b, coef=si_h_s, resp=ca),
  prior(normal(0.33,0.19), class=b, coef=mor_h_s, resp=ca),
  prior(normal(0.35,0.18), class=b, coef=ce_h_s, resp=ca),
  prior(normal(0.35,0.2), class=b, coef=an_h_s, resp=ca),
  #for the mediation models
  prior(normal(0,0.5), class=b, resp=an),
  prior(normal(0,0.5), class=b, resp=ce))

c4_m2 <- brm(data=d_long_c4_m2,
             family=cumulative(),
             main_model + mediation_an + mediation_ce,
             iter=100, warmup=20, chains=2,
             prior=prior_m2,
             seed=1)

summary(c4_m2)
```

See if quicker

```{r}
main_model <- bf(ca ~ 
                   si_h_s + mor_h_s + ce_h_s + an_h_s +
                   house+(1|ID)+item_ca)

mediation_an <- bf(an ~ si_h_s + mor_h_s + (1|ID) + item_an)
mediation_ce <- bf(ce ~ si_h_s + mor_h_s + (1|ID) + item_ce)

prior_m2 <- c(
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=ca),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=ca),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=ca),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=ca),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=ca),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=ca),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=an),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=an),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=an),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=an),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=an),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=an),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=ce),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=ce),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=ce),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=ce),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=ce),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=ce),
  #random intercepts => so they can vary across 0
  prior(normal(0, 1), class=sd, resp=ca),
  prior(normal(0, 1), class=sd, resp=an),
  prior(normal(0, 1), class=sd, resp=ce),
  #predictors
  prior(normal(0.37,0.2), class=b, coef=si_h_s, resp=ca),
  prior(normal(0.33,0.19), class=b, coef=mor_h_s, resp=ca),
  prior(normal(0.35,0.18), class=b, coef=ce_h_s, resp=ca),
  prior(normal(0.35,0.2), class=b, coef=an_h_s, resp=ca),
  #for the mediation models
  prior(normal(0,0.5), class=b, resp=an),
  prior(normal(0,0.5), class=b, resp=ce))

c4_m2 <- brm(data=d_long_c4_m2,
             family=cumulative(),
             main_model + mediation_an + mediation_ce,
             iter=500, warmup=40, chains=2,
             prior=prior_m2,
             seed=1)

summary(c4_m2)
```

## Model comparison

```{r}
c5_empty <- readRDS("fits/c5_empty.rds")
c5_m0 <- readRDS("fits/c5_m0.rds")
c5_m1 <- readRDS("fits/c5_m1.rds")

(loo(c5_empty,c5_m0,c5_m1))$diffs %>% 
  as.data.frame() %>% 
  mutate(ratio=elpd_diff/se_diff)

summary(c5_empty)
summary(c5_m0)
summary(c5_m1)
```
