---
title: "Models motivations to join"
output: html_document
---

```{r}
library(tidyverse)
library(brms)
library(patchwork)
```

```{r}
d <- read.csv("data/d_scales.csv")
```

# Economic motivation to join

## Prior predictive check

```{r}
prior_empty <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6))

#logit empty model
eco_ppc_logit <- brm(data=d,
      family=cumulative(),
      prior=prior_empty,
      dv_market_1 ~ 1 + (1|ID),
      iter=4000, warmup=1000, chains=4, cores=4,
      sample_prior = "only",
      seed=1)

#probit empty model
eco_ppc_probit <- brm(data=d,
      family=cumulative(link=probit),
      prior=prior_empty,
      dv_market_1 ~ 1 + (1|ID),
      iter=4000, warmup=1000, chains=4, cores=4,
      sample_prior = "only",
      seed=1)

plot_eco_logit <- pp_check(eco_ppc_logit, type="bars") +
  ggtitle("Logit link function")
plot_eco_probit <- pp_check(eco_ppc_probit, type="bars") +
  ggtitle("Probit link function")

plot_eco_logit|plot_eco_probit

#model comparison
eco_logit <- brm(data=d,
      family=cumulative(),
      prior=prior_empty,
      dv_market_1 ~ 1 + (1|ID),
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
      seed=1)

eco_probit <- brm(data=d,
      family=cumulative(link=probit),
      prior=prior_empty,
      dv_market_1 ~ 1 + (1|ID),
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
      seed=1)

saveRDS(eco_logit, "fit/eco_logit.rds")
saveRDS(eco_probit, "fit/eco_probit.rds")

loo(eco_logit,eco_probit)

eco_empty <- readRDS("fit/eco_logit.rds")
```

Probit seems better, but at the same time it has a lot of estimation problems, so keep logit

## Adding `(1|house)`

```{r}
prior_house <- c(prior(normal(-1.07,2), class=Intercept, coef=1),
                 prior(normal(-0.57,2), class=Intercept, coef=2),
                 prior(normal(-0.18,2), class=Intercept, coef=3),
                 prior(normal(0.18,2), class=Intercept, coef=4),
                 prior(normal(0.57,2), class=Intercept, coef=5),
                 prior(normal(1.07,2), class=Intercept, coef=6),
                 prior(exponential(1), class=sd))

eco_house <- brm(data=d,
      family=cumulative(),
      prior=prior_house,
      dv_market_1 ~ (1|house) + (1|ID),
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(8),
      backend="cmdstanr", silent=0,
      seed=1)
saveRDS(eco_house, "fit/eco_house.rds")

summary(eco_house)
loo(eco_empty,eco_house)
```

## Adding the socio-demographic variables

```{r}
prior_sd <- c(prior(normal(-1.07,2), class=Intercept, coef=1),
                 prior(normal(-0.57,2), class=Intercept, coef=2),
                 prior(normal(-0.18,2), class=Intercept, coef=3),
                 prior(normal(0.18,2), class=Intercept, coef=4),
                 prior(normal(0.57,2), class=Intercept, coef=5),
                 prior(normal(1.07,2), class=Intercept, coef=6),
                 prior(exponential(1), class=sd),
                 prior(normal(0,0.5), class=b))

eco_sd <- brm(data=d,
      family=cumulative(),
      prior=prior_sd,
      dv_market_1 ~ (1|house) + (1|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation),
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(8),
      backend="cmdstanr", silent=0,
      seed=1)
saveRDS(eco_sd, "fit/eco_sd.rds")

summary(eco_sd)
loo(eco_house,eco_sd)
```

## Adding the predictors

```{r}
prior_pred <- c(prior(normal(-1.07,2), class=Intercept, coef=1),
                 prior(normal(-0.57,2), class=Intercept, coef=2),
                 prior(normal(-0.18,2), class=Intercept, coef=3),
                 prior(normal(0.18,2), class=Intercept, coef=4),
                 prior(normal(0.57,2), class=Intercept, coef=5),
                 prior(normal(1.07,2), class=Intercept, coef=6),
                 prior(exponential(1), class=sd),
                 #beta coefficients = no expectations since sd variables
              prior(normal(0,0.5), class=b, coef=gender_m_w),
              prior(normal(0,0.5), class=b, coef=age_s),
              prior(normal(0,0.5), class=b, coef=income_s),
              prior(normal(0,0.5), class=b, coef=mopol_orientation),
              prior(normal(0,0.5), class=b, coef=moeducation),
              #priors for the predictors, calculated based on the results of the meta-analysis
              prior(normal(0.37,0.2), class=b, coef=si_m_e_s),
              prior(normal(0.33,0.19), class=b, coef=mor_m_s),
              prior(normal(0.35,0.18), class=b, coef=ce_m_s),
              prior(normal(0.35,0.2), class=b, coef=an_m_eco_s))

eco_pred <- brm(data=d,
      family=cumulative(),
      prior=prior_pred,
      dv_market_1 ~ (1|house) + (1|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + ce_m_s + an_m_eco_s,
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
      backend="cmdstanr", silent=0,
      seed=1)
saveRDS(eco_pred_id, "fit/eco_pred.rds")

summary(eco_pred)
loo(eco_sd,eco_pred)

#test using the dataset below to be able to compare
eco_pred_test <- brm(data=d_eco_full,
      family=cumulative(),
      prior=prior_pred,
      dv_market_1 ~ (1|house) + (1|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + ce_m_s + an_m_eco_s,
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(8),
      backend = "cmdstanr",
      seed=1)

summary(eco_pred_test)
```

## Full model

```{r}
d_eco_full <- d %>% 
  pivot_longer(cols=c("ce_market_1","ce_market_2")) %>% 
  rename(answer_ce=value,
         item_ce=name) %>% 
  pivot_longer(cols=starts_with("anger_market2")) %>% 
  rename(answer_an=value,
         item_an=name)
```

```{r}
MAIN <- bf(dv_market_1 ~ (1|h|house) + (1|i|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + ce_m_s + an_m_eco_s)

CE <- bf(answer_ce ~ (1|h|house) + (1|i|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + item_ce)
  
AN <- bf(answer_an ~ (1|h|house) + (1|i|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + item_an)

get_prior(data=d_eco_full,
      family=cumulative(),
      MAIN + CE + AN)

prior_full <- c(
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=dvmarket1),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=dvmarket1),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=dvmarket1),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=dvmarket1),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=dvmarket1),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=dvmarket1),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=answeran),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=answeran),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=answeran),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=answeran),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=answeran),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=answeran),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=answerce),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=answerce),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=answerce),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=answerce),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=answerce),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=answerce),
  #random intercepts => so they can vary across 0
  prior(normal(0, 1), class=sd, resp=dvmarket1),
  prior(normal(0, 1), class=sd, resp=answeran),
  prior(normal(0, 1), class=sd, resp=answerce),
  #beta coefficients = no expectations since sd variables
  prior(normal(0,0.5), class=b, coef=gender_m_w, resp=dvmarket1),
  prior(normal(0,0.5), class=b, coef=age_s, resp=dvmarket1),
  prior(normal(0,0.5), class=b, coef=income_s, resp=dvmarket1),
  prior(normal(0,0.5), class=b, coef=mopol_orientation, resp=dvmarket1),
  prior(normal(0,0.5), class=b, coef=moeducation, resp=dvmarket1),
  #predictors
  prior(normal(0.37,0.2), class=b, coef=si_m_e_s, resp=dvmarket1),
  prior(normal(0.33,0.19), class=b, coef=mor_m_s, resp=dvmarket1),
  prior(normal(0.35,0.18), class=b, coef=ce_m_s, resp=dvmarket1),
  prior(normal(0.35,0.2), class=b, coef=an_m_eco_s, resp=dvmarket1),
  #for the mediation models
  prior(normal(0,0.5), class=b, resp=answeran),
  prior(normal(0,0.5), class=b, resp=answerce))

eco_full <- brm(data=d_eco_full,
      family=cumulative(),
      prior=prior_full,
      MAIN + CE + AN,
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(8),
      backend = "cmdstanr",
      seed=1)

summary(eco_full)
saveRDS(eco_full, "fit/eco_full.rds")
eco_full_old <- readRDS("fit/eco_full_old.rds") #this is the model without allowing house and ID to covariate across equations => clearly worse than the new version, see loo
eco_full <- readRDS("fit/eco_full.rds")
loo(eco_full,eco_full_old)
```

## Comparison of all models

```{r}
eco_empty <- readRDS("fit/eco_logit.rds")
eco_house <- readRDS("fit/eco_house.rds")
eco_sd <- readRDS("fit/eco_sd.rds")
eco_pred <- readRDS("fit/eco_pred.rds")
eco_full <- readRDS("fit/eco_full.rds")

loo(eco_empty,
    eco_house,
    eco_sd,
    eco_pred)
```

# Social motivation

```{r}
d_soc <- d %>% 
  pivot_longer(cols=c("dv_market_2","dv_market_3")) %>% 
  rename(answer=value,
         item=name)
```

## PPC

```{r}
prior_empty <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6),
              prior(exponential(1), class=sd),
              prior(normal(0,0.5), class=b))

prior_empty_dT <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1, group="dv_market_2"),
              prior(normal(-0.57,2), class=Intercept, coef=2, group="dv_market_2"),
              prior(normal(-0.18,2), class=Intercept, coef=3, group="dv_market_2"),
              prior(normal(0.18,2), class=Intercept, coef=4, group="dv_market_2"),
              prior(normal(0.57,2), class=Intercept, coef=5, group="dv_market_2"),
              prior(normal(1.07,2), class=Intercept, coef=6, group="dv_market_2"),
              prior(normal(-1.07,2), class=Intercept, coef=1, group="dv_market_3"),
              prior(normal(-0.57,2), class=Intercept, coef=2, group="dv_market_3"),
              prior(normal(-0.18,2), class=Intercept, coef=3, group="dv_market_3"),
              prior(normal(0.18,2), class=Intercept, coef=4, group="dv_market_3"),
              prior(normal(0.57,2), class=Intercept, coef=5, group="dv_market_3"),
              prior(normal(1.07,2), class=Intercept, coef=6, group="dv_market_3"),
              prior(exponential(1), class=sd),
              prior(normal(0,0.5), class=b))

soc_ppc_logit_sT <- brm(data=d_soc,
      family=cumulative(),
      prior=prior_empty,
      answer ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      sample_prior = "only",
      seed=1)

soc_ppc_logit_dT <- brm(data=d_soc,
      family=cumulative(),
      prior=prior_empty_dT,
      answer | thres(gr = item) ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      sample_prior = "only",
      seed=1)

soc_ppc_probit_sT <- brm(data=d_soc,
      family=cumulative(link=probit),
      prior=prior_empty,
      answer ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      sample_prior = "only",
      seed=1)

soc_ppc_probit_dT <- brm(data=d_soc,
      family=cumulative(link=probit),
      prior=prior_empty_dT,
      answer | thres(gr = item) ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      sample_prior = "only",
      seed=1)

plot_soc_logit_sT <- pp_check(soc_ppc_logit_sT, type="bars") +
  ggtitle("Logit, same T")
plot_soc_logit_dT <- pp_check(soc_ppc_logit_dT, type="bars") +
  ggtitle("Logit, diff T")
plot_soc_probit_sT <- pp_check(soc_ppc_probit_sT, type="bars") +
  ggtitle("Probit, same T")
plot_soc_probit_dT <- pp_check(soc_ppc_probit_dT, type="bars") +
  ggtitle("Probit, diff T")

(plot_soc_logit_sT|plot_soc_logit_dT) /
  (plot_soc_probit_sT|plot_soc_probit_dT)
```

```{r}
#model comparison
soc_logit_sT <- brm(data=d_soc,
      family=cumulative(),
      prior=prior_empty,
      answer ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(soc_logit_sT,"fit/soc_logit_sT.rds")

soc_logit_dT <- brm(data=d_soc,
      family=cumulative(),
      prior=prior_empty_dT,
      answer | thres(gr = item) ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(soc_logit_dT,"fit/soc_logit_dT.rds")

soc_probit_sT <- brm(data=d_soc,
      family=cumulative(link=probit),
      prior=prior_empty,
      answer ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(soc_probit_sT,"fit/soc_probit_sT.rds")

soc_probit_dT <- brm(data=d_soc,
      family=cumulative(link=probit),
      prior=prior_empty_dT,
      answer | thres(gr = item) ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(soc_probit_sT,"fit/soc_probit_dT.rds")

loo(soc_logit_sT,
    soc_logit_dT,
    soc_probit_sT,
    soc_probit_dT)

soc_empty <- readRDS("fit/soc_logit_sT.rds")
```

`soc_probit_sT` seems to be the better model, but it has a lot of `Pareto_k >0.7`, while `soc_logit_sT` is not significantly worse and does not have the `Pareto_k` problem =\> keep logit_sT

## Adding `(1|house)`

```{r}
prior_house <- c(prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6),
              prior(exponential(1), class=sd),
              prior(normal(0,0.5), class=b))

soc_house <- brm(data=d_soc,
      family=cumulative(),
      prior=prior_house,
      answer ~ (1|house) + (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(soc_house,"fit/soc_house.rds")

summary(soc_house)
loo(soc_empty,soc_house)
```

## Adding the socio-demographic

```{r}
prior_sd <- c(prior(normal(-1.07,2), class=Intercept, coef=1),
                 prior(normal(-0.57,2), class=Intercept, coef=2),
                 prior(normal(-0.18,2), class=Intercept, coef=3),
                 prior(normal(0.18,2), class=Intercept, coef=4),
                 prior(normal(0.57,2), class=Intercept, coef=5),
                 prior(normal(1.07,2), class=Intercept, coef=6),
                 prior(exponential(1), class=sd),
                 prior(normal(0,0.5), class=b))

soc_sd <- brm(data=d_soc,
      family=cumulative(),
      prior=prior_sd,
      answer ~ (1|house) + (1|ID) + item +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation),
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(soc_sd,"fit/soc_sd.rds")

summary(soc_sd)
loo(soc_house,soc_sd)
```

## Adding the predictors

```{r}
prior_pred <- c(prior(normal(-1.07,2), class=Intercept, coef=1),
                 prior(normal(-0.57,2), class=Intercept, coef=2),
                 prior(normal(-0.18,2), class=Intercept, coef=3),
                 prior(normal(0.18,2), class=Intercept, coef=4),
                 prior(normal(0.57,2), class=Intercept, coef=5),
                 prior(normal(1.07,2), class=Intercept, coef=6),
                 prior(exponential(1), class=sd),
                 #beta coefficients = no expectations since sd variables
              prior(normal(0,0.5), class=b, coef=gender_m_w),
              prior(normal(0,0.5), class=b, coef=age_s),
              prior(normal(0,0.5), class=b, coef=income_s),
              prior(normal(0,0.5), class=b, coef=mopol_orientation),
              prior(normal(0,0.5), class=b, coef=moeducation),
              #priors for the predictors, calculated based on the results of the meta-analysis
              prior(normal(0.37,0.2), class=b, coef=si_m_e_s),
              prior(normal(0.33,0.19), class=b, coef=mor_m_s),
              prior(normal(0.35,0.18), class=b, coef=ce_m_s),
              prior(normal(0.35,0.2), class=b, coef=an_m_soc_s))

soc_pred <- brm(data=d_soc,
      family=cumulative(),
      prior=prior_pred,
      answer ~ (1|house) + (1|ID) + item +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + ce_m_s + an_m_soc_s,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(soc_pred,"fit/soc_pred.rds")

summary(soc_pred)
loo(soc_pred,soc_sd)
```

## Full model

```{r}
d_soc_full <- d_soc %>% 
  rename(answer_ca=answer,
         item_ca=item) %>% 
  pivot_longer(cols=c("ce_market_1","ce_market_2")) %>% 
  rename(answer_ce=value,
         item_ce=name) %>% 
  pivot_longer(cols=starts_with("anger_market1")) %>% 
  rename(answer_an=value,
         item_an=name)
```

```{r}
MAIN <- bf(answer_ca ~ (1|h|house) + (1|i|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + ce_m_s + an_m_eco_s + item_ca)

CE <- bf(answer_ce ~ (1|h|house) + (1|i|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + item_ce)
  
AN <- bf(answer_an ~ (1|h|house) + (1|i|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + item_an)

get_prior(data=d_eco_full,
      family=cumulative(),
      MAIN + CE + AN)

prior_full <- c(
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=answerca),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=answerca),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=answerca),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=answerca),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=answerca),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=answerca),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=answeran),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=answeran),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=answeran),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=answeran),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=answeran),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=answeran),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=answerce),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=answerce),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=answerce),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=answerce),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=answerce),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=answerce),
  #random intercepts => so they can vary across 0
  prior(normal(0, 1), class=sd, resp=answerca),
  prior(normal(0, 1), class=sd, resp=answeran),
  prior(normal(0, 1), class=sd, resp=answerce),
  #beta coefficients = no expectations since sd variables
  prior(normal(0,0.5), class=b, coef=gender_m_w, resp=answerca),
  prior(normal(0,0.5), class=b, coef=age_s, resp=answerca),
  prior(normal(0,0.5), class=b, coef=income_s, resp=answerca),
  prior(normal(0,0.5), class=b, coef=mopol_orientation, resp=answerca),
  prior(normal(0,0.5), class=b, coef=moeducation, resp=answerca),
  #predictors
  prior(normal(0.37,0.2), class=b, coef=si_m_e_s, resp=answerca),
  prior(normal(0.33,0.19), class=b, coef=mor_m_s, resp=answerca),
  prior(normal(0.35,0.18), class=b, coef=ce_m_s, resp=answerca),
  prior(normal(0.35,0.2), class=b, coef=an_m_eco_s, resp=answerca),
  #for the mediation models
  prior(normal(0,0.5), class=b, resp=answeran),
  prior(normal(0,0.5), class=b, resp=answerce))

soc_full <- brm(data=d_soc_full,
      family=cumulative(),
      prior=prior_full,
      MAIN + CE + AN,
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(8),
      backend="cmdstanr", silent=1,
      seed=1)

summary(soc_full)
saveRDS(soc_full, "fit/soc_full.rds")

soc_full_old <- readRDS("fit/soc_full_old.rds") #the one without letting the house and id levels correlate among equations
soc_full <- readRDS("fit/soc_full.rds")

loo(soc_full,soc_full_old)
```

# Environmental motivation

```{r}
d_env <- d %>% 
  pivot_longer(cols=starts_with("dv_env_")) %>% 
  rename(answer=value,
         item=name)
```

## PPC

```{r}
prior_empty <- c(#first item
              prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6),
              prior(exponential(1), class=sd),
              prior(normal(0,0.5), class=b))

prior_empty_dT <- c(
  prior(normal(-1.07,2), class=Intercept, coef=1, group="dv_env_1"),
              prior(normal(-0.57,2), class=Intercept, coef=2, group="dv_env_1"),
              prior(normal(-0.18,2), class=Intercept, coef=3, group="dv_env_1"),
              prior(normal(0.18,2), class=Intercept, coef=4, group="dv_env_1"),
              prior(normal(0.57,2), class=Intercept, coef=5, group="dv_env_1"),
              prior(normal(1.07,2), class=Intercept, coef=6, group="dv_env_1"),
              prior(normal(-1.07,2), class=Intercept, coef=1, group="dv_env_2"),
              prior(normal(-0.57,2), class=Intercept, coef=2, group="dv_env_2"),
              prior(normal(-0.18,2), class=Intercept, coef=3, group="dv_env_2"),
              prior(normal(0.18,2), class=Intercept, coef=4, group="dv_env_2"),
              prior(normal(0.57,2), class=Intercept, coef=5, group="dv_env_2"),
              prior(normal(1.07,2), class=Intercept, coef=6, group="dv_env_2"),
              prior(normal(-1.07,2), class=Intercept, coef=1, group="dv_env_3"),
              prior(normal(-0.57,2), class=Intercept, coef=2, group="dv_env_3"),
              prior(normal(-0.18,2), class=Intercept, coef=3, group="dv_env_3"),
              prior(normal(0.18,2), class=Intercept, coef=4, group="dv_env_3"),
              prior(normal(0.57,2), class=Intercept, coef=5, group="dv_env_3"),
              prior(normal(1.07,2), class=Intercept, coef=6, group="dv_env_3"),
              prior(exponential(1), class=sd),
              prior(normal(0,0.5), class=b))

env_ppc_logit_sT <- brm(data=d_env,
      family=cumulative(),
      prior=prior_empty,
      answer ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      sample_prior = "only",
      seed=1)

env_ppc_logit_dT <- brm(data=d_env,
      family=cumulative(),
      prior=prior_empty_dT,
      answer | thres(gr = item) ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      sample_prior = "only",
      seed=1)

env_ppc_probit_sT <- brm(data=d_env,
      family=cumulative(link=probit),
      prior=prior_empty,
      answer ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      sample_prior = "only",
      seed=1)

env_ppc_probit_dT <- brm(data=d_env,
      family=cumulative(link=probit),
      prior=prior_empty_dT,
      answer | thres(gr = item) ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      sample_prior = "only",
      seed=1)

plot_env_logit_sT <- pp_check(env_ppc_logit_sT, type="bars") +
  ggtitle("Logit, same T")
plot_env_logit_dT <- pp_check(env_ppc_logit_dT, type="bars") +
  ggtitle("Logit, diff T")
plot_env_probit_sT <- pp_check(env_ppc_probit_sT, type="bars") +
  ggtitle("Probit, same T")
plot_env_probit_dT <- pp_check(env_ppc_probit_dT, type="bars") +
  ggtitle("Probit, diff T")

(plot_env_logit_sT|plot_env_logit_dT) /
  (plot_env_probit_sT|plot_env_probit_dT)
```

```{r}
#model comparison
env_logit_sT <- brm(data=d_env,
      family=cumulative(),
      prior=prior_empty,
      answer ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(env_logit_sT,"fit/env_logit_sT.rds")

env_logit_dT <- brm(data=d_env,
      family=cumulative(),
      prior=prior_empty_dT,
      answer | thres(gr = item) ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(env_logit_dT,"fit/env_logit_dT.rds")

env_probit_sT <- brm(data=d_env,
      family=cumulative(link=probit),
      prior=prior_empty,
      answer ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(env_probit_sT,"fit/env_probit_sT.rds")

env_probit_dT <- brm(data=d_env,
      family=cumulative(link=probit),
      prior=prior_empty_dT,
      answer | thres(gr = item) ~ (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(env_probit_sT,"fit/env_probit_dT.rds")

loo(env_logit_sT,
    env_logit_dT,
    env_probit_sT,
    env_probit_dT)

env_empty <- readRDS("fit/env_logit_sT.rds")
```

`env_logit_sT` is clearly the best model

## Adding `(1|house)`

```{r}
prior_house <- c(prior(normal(-1.07,2), class=Intercept, coef=1),
              prior(normal(-0.57,2), class=Intercept, coef=2),
              prior(normal(-0.18,2), class=Intercept, coef=3),
              prior(normal(0.18,2), class=Intercept, coef=4),
              prior(normal(0.57,2), class=Intercept, coef=5),
              prior(normal(1.07,2), class=Intercept, coef=6),
              prior(exponential(1), class=sd),
              prior(normal(0,0.5), class=b))

env_house <- brm(data=d_env,
      family=cumulative(),
      prior=prior_house,
      answer ~ (1|house) + (1|ID) + item,
      iter=4000, warmup=1000, chains=4, cores=4,
      seed=1)
saveRDS(env_house,"fit/env_house.rds")

summary(env_house)
loo(env_empty,env_house)
```

## Adding the socio-demographic

```{r}
prior_sd <- c(prior(normal(-1.07,2), class=Intercept, coef=1),
                 prior(normal(-0.57,2), class=Intercept, coef=2),
                 prior(normal(-0.18,2), class=Intercept, coef=3),
                 prior(normal(0.18,2), class=Intercept, coef=4),
                 prior(normal(0.57,2), class=Intercept, coef=5),
                 prior(normal(1.07,2), class=Intercept, coef=6),
                 prior(exponential(1), class=sd),
                 prior(normal(0,0.5), class=b)) 

env_sd <- brm(data=d_env,
      family=cumulative(),
      prior=prior_sd,
      answer ~ (1|house) + (1|ID) + item +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation),
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(5),
      seed=1)
saveRDS(env_sd,"fit/env_sd.rds") 
 
summary(env_sd) 
loo(env_house,env_sd) 
```

## Adding the predictors

```{r}
prior_pred <- c(prior(normal(-1.07,2), class=Intercept, coef=1),
                 prior(normal(-0.57,2), class=Intercept, coef=2),
                 prior(normal(-0.18,2), class=Intercept, coef=3),
                 prior(normal(0.18,2), class=Intercept, coef=4),
                 prior(normal(0.57,2), class=Intercept, coef=5),
                 prior(normal(1.07,2), class=Intercept, coef=6),
                 prior(exponential(1), class=sd),
                 #beta coefficients = no expectations since sd variables
              prior(normal(0,0.5), class=b, coef=gender_m_w),
              prior(normal(0,0.5), class=b, coef=age_s),
              prior(normal(0,0.5), class=b, coef=income_s),
              prior(normal(0,0.5), class=b, coef=mopol_orientation),
              prior(normal(0,0.5), class=b, coef=moeducation),
              #priors for the predictors, calculated based on the results of the meta-analysis
              prior(normal(0.37,0.2), class=b, coef=si_m_e_s),
              prior(normal(0.33,0.19), class=b, coef=mor_e_s),
              prior(normal(0.35,0.18), class=b, coef=ce_e_s),
              prior(normal(0.35,0.2), class=b, coef=an_e_s))

env_pred <- brm(data=d_env,
      family=cumulative(),
      prior=prior_pred,
      answer ~ (1|house) + (1|ID) + item +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_e_s + ce_e_s + an_e_s,
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(5),
      seed=1)
saveRDS(env_pred,"fit/env_pred.rds")

summary(env_pred)
loo(env_pred,env_sd)
```

## Full model

```{r}
d_env_full <- d_env %>% 
  rename(answer_ca=answer,
         item_ca=item) %>% 
  pivot_longer(cols=c("ce_env_1","ce_env_2")) %>% 
  rename(answer_ce=value,
         item_ce=name) %>% 
  pivot_longer(cols=starts_with("anger_env")) %>% 
  rename(answer_an=value,
         item_an=name)
```

```{r}
MAIN <- bf(answer_ca ~ (1|h|house) + (1|i|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_e_s + ce_e_s + an_e_s + item_ca)

CE <- bf(answer_ce ~ (1|h|house) + (1|i|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_e_s + item_ce)
  
AN <- bf(answer_an ~ (1|h|house) + (1|i|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_e_s + item_an)

get_prior(data=d_env_full,
      family=cumulative(),
      MAIN + CE + AN)

prior_full <- c(
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=answerca),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=answerca),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=answerca),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=answerca),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=answerca),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=answerca),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=answeran),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=answeran),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=answeran),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=answeran),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=answeran),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=answeran),
  prior(normal(-1.07,2), class=Intercept, coef=1, resp=answerce),
  prior(normal(-0.57,2), class=Intercept, coef=2, resp=answerce),
  prior(normal(-0.18,2), class=Intercept, coef=3, resp=answerce),
  prior(normal(0.18,2), class=Intercept, coef=4, resp=answerce),
  prior(normal(0.57,2), class=Intercept, coef=5, resp=answerce),
  prior(normal(1.07,2), class=Intercept, coef=6, resp=answerce),
  #random intercepts => so they can vary across 0
  prior(normal(0, 1), class=sd, resp=answerca),
  prior(normal(0, 1), class=sd, resp=answeran),
  prior(normal(0, 1), class=sd, resp=answerce),
  #beta coefficients = no expectations since sd variables
  prior(normal(0,0.5), class=b, coef=gender_m_w, resp=answerca),
  prior(normal(0,0.5), class=b, coef=age_s, resp=answerca),
  prior(normal(0,0.5), class=b, coef=income_s, resp=answerca),
  prior(normal(0,0.5), class=b, coef=mopol_orientation, resp=answerca),
  prior(normal(0,0.5), class=b, coef=moeducation, resp=answerca),
  #predictors
  prior(normal(0.37,0.2), class=b, coef=si_m_e_s, resp=answerca),
  prior(normal(0.33,0.19), class=b, coef=mor_e_s, resp=answerca),
  prior(normal(0.35,0.18), class=b, coef=ce_e_s, resp=answerca),
  prior(normal(0.35,0.2), class=b, coef=an_e_s, resp=answerca),
  #for the mediation models
  prior(normal(0,0.5), class=b, resp=answeran),
  prior(normal(0,0.5), class=b, resp=answerce))

env_full <- brm(data=d_env_full,
      family=cumulative(),
      prior=prior_full,
      MAIN + CE + AN,
      iter=4000, warmup=1000, chains=4, cores=4, threads=threading(8),
      backend="cmdstanr", silent=0,
      seed=1)

summary(env_full)
saveRDS(env_full, "fit/env_full.rds")

env_full_old <- readRDS("fit/env_full_old.rds")
env_full <- readRDS("fit/env_full.rds")

loo(env_full, env_full_old)
```
