---
title: "Supplementary materials - Cross sectional analyses about the motivations to
  join"
output: html_document
---

```{r}
library(tidyverse)
library(brms)
```

```{r}
d <- read.csv("data/d_scales.csv")
```

# 1. Economic motivation

## M0: socio-demographic

```{r}
prior_m0 <-    c(prior(normal(-1.79,2), class=Intercept, coef=1),
                 prior(normal(-0.92,2), class=Intercept, coef=2),
                 prior(normal(-0.29,2), class=Intercept, coef=3),
                 prior(normal(0.29,2), class=Intercept, coef=4),
                 prior(normal(0.92,2), class=Intercept, coef=5),
                 prior(normal(1.79,2), class=Intercept, coef=6),
                 prior(exponential(1), class=sd),
                 prior(normal(0,0.5), class=b))

eco_m0 <- brm(data=d,
              family=cumulative(),
              prior=prior_sd,
              dv_market_1 ~ (1|house) + (1|ID) +
                gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation),
              iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
              backend="cmdstanr", silent=0,
              seed=1)

saveRDS(eco_m0, "fit/eco_m0.rds")
```

## M1: predictors

```{r}
prior_m1 <- c(prior(normal(-1.79,2), class=Intercept, coef=1),
              prior(normal(-0.92,2), class=Intercept, coef=2),
              prior(normal(-0.29,2), class=Intercept, coef=3),
              prior(normal(0.29,2), class=Intercept, coef=4),
              prior(normal(0.92,2), class=Intercept, coef=5),
              prior(normal(1.79,2), class=Intercept, coef=6),
              prior(exponential(1), class=sd),
              #beta coefficients = no expectations since sd variables
              prior(normal(0,0.5), class=b, coef=gender_m_w),
              prior(normal(0,0.5), class=b, coef=age_s),
              prior(normal(0,0.5), class=b, coef=income_s),
              prior(normal(0,0.5), class=b, coef=mopol_orientation),
              prior(normal(0,0.5), class=b, coef=moeducation),
              #priors for the predictors, calculated based on the results of the meta-analysis
              prior(normal(0.37,0.2), class=b, coef=si_m_e_s),
              prior(normal(0.33,0.19), class=b, coef=mor_m_s),
              prior(normal(0.35,0.18), class=b, coef=ce_m_s),
              prior(normal(0.35,0.2), class=b, coef=an_m_eco_s))

eco_m1 <- brm(data=d,
              family=cumulative(),
              prior=prior_pred,
              dv_market_1 ~ (1|house) + (1|ID) +
                gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
                si_m_e_s + mor_m_s + ce_m_s + an_m_eco_s,
              iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
              backend="cmdstanr", silent=0,
              seed=1)

saveRDS(eco_m1, "fit/eco_m1.rds")
```

## M2: indirect paths (coll. efficacy and anger predicted by soc. identity and morality)

```{r}
d_eco_ce_an <- d %>% 
  pivot_longer(cols=c("ce_market_1","ce_market_2")) %>% 
  rename(answer_ce=value,
         item_ce=name) %>% 
  pivot_longer(cols=starts_with("anger_market2")) %>% 
  rename(answer_an=value,
         item_an=name)

prior_m2 <- c(prior(normal(-1.79,2), class=Intercept, coef=1, resp=answeran),
              prior(normal(-0.92,2), class=Intercept, coef=2, resp=answeran),
              prior(normal(-0.29,2), class=Intercept, coef=3, resp=answeran),
              prior(normal(0.29,2), class=Intercept, coef=4, resp=answeran),
              prior(normal(0.92,2), class=Intercept, coef=5, resp=answeran),
              prior(normal(1.79,2), class=Intercept, coef=6, resp=answeran),
              prior(normal(-1.79,2), class=Intercept, coef=1, resp=answerce),
              prior(normal(-0.92,2), class=Intercept, coef=2, resp=answerce),
              prior(normal(-0.29,2), class=Intercept, coef=3, resp=answerce),
              prior(normal(0.29,2), class=Intercept, coef=4, resp=answerce),
              prior(normal(0.92,2), class=Intercept, coef=5, resp=answerce),
              prior(normal(1.79,2), class=Intercept, coef=6, resp=answerce),
              #random intercepts => so they can vary across 0
              prior(normal(0,1), class=sd, resp=answeran),
              prior(normal(0,1), class=sd, resp=answerce),
              prior(normal(0,0.5), class=b, resp=answeran),
              prior(normal(0,0.5), class=b, resp=answerce))

CE <- bf(answer_ce ~ (1|house) + (1|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + item_ce)
  
AN <- bf(answer_an ~ (1|house) + (1|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + item_an)

eco_m2 <- brm(data=d_eco_ce_an,
              family=cumulative(),
              prior=prior_ind,
              CE + AN,
              iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
              backend = "cmdstanr",
              seed=1)

saveRDS(eco_m2, "fit/eco_m2.rds")
```

# 2. Social motivation

```{r}
d_soc <- d %>% 
  pivot_longer(cols=c("dv_market_2","dv_market_3")) %>% 
  rename(answer=value,
         item=name)
```

## Test same vs diff tresholds

```{r}
prior_sameT <- c(prior(normal(-1.79,2), class=Intercept, coef=1),
                 prior(normal(-0.92,2), class=Intercept, coef=2),
                 prior(normal(-0.29,2), class=Intercept, coef=3),
                 prior(normal(0.29,2), class=Intercept, coef=4),
                 prior(normal(0.92,2), class=Intercept, coef=5),
                 prior(normal(1.79,2), class=Intercept, coef=6),
                 prior(exponential(1), class=sd),
                 prior(normal(0,0.5), class=b))

prior_diffT <- c(prior(normal(-1.79,2), class=Intercept, coef=1, group="dv_market_2"),
                 prior(normal(-0.92,2), class=Intercept, coef=2, group="dv_market_2"),
                 prior(normal(-0.29,2), class=Intercept, coef=3, group="dv_market_2"),
                 prior(normal(0.29,2), class=Intercept, coef=4, group="dv_market_2"),
                 prior(normal(0.92,2), class=Intercept, coef=5, group="dv_market_2"),
                 prior(normal(1.79,2), class=Intercept, coef=6, group="dv_market_2"),
                 prior(normal(-1.79,2), class=Intercept, coef=1, group="dv_market_3"),
                 prior(normal(-0.92,2), class=Intercept, coef=2, group="dv_market_3"),
                 prior(normal(-0.29,2), class=Intercept, coef=3, group="dv_market_3"),
                 prior(normal(0.29,2), class=Intercept, coef=4, group="dv_market_3"),
                 prior(normal(0.92,2), class=Intercept, coef=5, group="dv_market_3"),
                 prior(normal(1.79,2), class=Intercept, coef=6, group="dv_market_3"),
                 prior(exponential(1), class=sd),
                 prior(normal(0,0.5), class=b))

soc_sameT <- brm(data=d_soc,
                 family=cumulative(),
                 prior=prior_sameT,
                 answer ~ (1|ID) + item,
                 iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
                 backend = "cmdstanr",
                 seed=1)
saveRDS(soc_sameT,"fit/soc_sameT.rds")

soc_diffT <- brm(data=d_soc,
                 family=cumulative(),
                 prior=prior_diffT,
                 answer | thres(gr = item) ~ (1|ID) + item,
                 iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
                 backend = "cmdstanr",
                 seed=1)
saveRDS(soc_diffT,"fit/soc_diffT.rds")

loo(soc_sameT,soc_diffT)
```

Basically the same, so keep the simplest one (same T)

## M0: socio-demographic

```{r}
soc_m0 <- brm(data=d_soc,
              family=cumulative(),
              prior=prior_m0,
              answer ~ (1|house) + (1|ID) + item +
                gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation),
              iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
              backend="cmdstanr", silent=1,
              seed=1)
saveRDS(soc_m0,"fit/soc_m0.rds")
```

## M1: predictors

```{r}
prior_m1 <- c(prior(normal(-1.79,2), class=Intercept, coef=1),
              prior(normal(-0.92,2), class=Intercept, coef=2),
              prior(normal(-0.29,2), class=Intercept, coef=3),
              prior(normal(0.29,2), class=Intercept, coef=4),
              prior(normal(0.92,2), class=Intercept, coef=5),
              prior(normal(1.79,2), class=Intercept, coef=6),
              prior(exponential(1), class=sd),
              #beta coefficients = no expectations since sd variables
              prior(normal(0,0.5), class=b, coef=gender_m_w),
              prior(normal(0,0.5), class=b, coef=age_s),
              prior(normal(0,0.5), class=b, coef=income_s),
              prior(normal(0,0.5), class=b, coef=mopol_orientation),
              prior(normal(0,0.5), class=b, coef=moeducation),
              #priors for the predictors, calculated based on the results of the meta-analysis
              prior(normal(0.37,0.2), class=b, coef=si_m_e_s),
              prior(normal(0.33,0.19), class=b, coef=mor_m_s),
              prior(normal(0.35,0.18), class=b, coef=ce_m_s),
              prior(normal(0.35,0.2), class=b, coef=an_m_soc_s))

soc_m1 <- brm(data=d_soc,
              family=cumulative(),
              prior=prior_m1,
              answer ~ (1|house) + (1|ID) + item +
                gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
                si_m_e_s + mor_m_s + ce_m_s + an_m_soc_s,
              iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
              backend="cmdstanr", silent=1,
              seed=1)
saveRDS(soc_m1,"fit/soc_m1.rds")
```

## M2: complete model

```{r}
d_soc_ce_an <- d %>% 
  pivot_longer(cols=c("ce_market_1","ce_market_2")) %>% 
  rename(answer_ce=value,
         item_ce=name) %>% 
  pivot_longer(cols=starts_with("anger_market1")) %>% 
  rename(answer_an=value,
         item_an=name)

CE <- bf(answer_ce ~ (1|house) + (1|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + item_ce)
  
AN <- bf(answer_an ~ (1|house) + (1|ID) +
        gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
        si_m_e_s + mor_m_s + item_an)

soc_m2 <- brm(data=d_soc_ce_an,
              family=cumulative(),
              prior=prior_m2,
              CE + AN,
              iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
              backend = "cmdstanr", silent=1,
              seed=1)

saveRDS(soc_m2, "fit/soc_m2.rds")
```

# 3. Environmental motivation

```{r}
d_env <- d %>% 
  pivot_longer(cols=starts_with("dv_env_")) %>% 
  rename(answer=value,
         item=name)
```

## Test same vs different thresholds

```{r}
prior_diffT <- c(prior(normal(-1.79,2), class=Intercept, coef=1, group="dv_env_1"),
                 prior(normal(-0.92,2), class=Intercept, coef=2, group="dv_env_1"),
                 prior(normal(-0.29,2), class=Intercept, coef=3, group="dv_env_1"),
                 prior(normal(0.29,2), class=Intercept, coef=4, group="dv_env_1"),
                 prior(normal(0.92,2), class=Intercept, coef=5, group="dv_env_1"),
                 prior(normal(1.79,2), class=Intercept, coef=6, group="dv_env_1"),
                 prior(normal(-1.79,2), class=Intercept, coef=1, group="dv_env_2"),
                 prior(normal(-0.92,2), class=Intercept, coef=2, group="dv_env_2"),
                 prior(normal(-0.29,2), class=Intercept, coef=3, group="dv_env_2"),
                 prior(normal(0.29,2), class=Intercept, coef=4, group="dv_env_2"),
                 prior(normal(0.92,2), class=Intercept, coef=5, group="dv_env_2"),
                 prior(normal(1.79,2), class=Intercept, coef=6, group="dv_env_2"),
                 prior(normal(-1.79,2), class=Intercept, coef=1, group="dv_env_3"),
                 prior(normal(-0.92,2), class=Intercept, coef=2, group="dv_env_3"),
                 prior(normal(-0.29,2), class=Intercept, coef=3, group="dv_env_3"),
                 prior(normal(0.29,2), class=Intercept, coef=4, group="dv_env_3"),
                 prior(normal(0.92,2), class=Intercept, coef=5, group="dv_env_3"),
                 prior(normal(1.79,2), class=Intercept, coef=6, group="dv_env_3"),
                 prior(exponential(1), class=sd),
                 prior(normal(0,0.5), class=b))

env_sameT <- brm(data=d_env,
                 family=cumulative(),
                 prior=prior_sameT,
                 answer ~ (1|ID) + item,
                 iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
                 backend = "cmdstanr",
                 seed=1)
saveRDS(env_sameT,"fit/env_sameT.rds")

env_diffT <- brm(data=d_env,
                 family=cumulative(),
                 prior=prior_diffT,
                 answer | thres(gr = item) ~ (1|ID) + item,
                 iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
                 backend = "cmdstanr",
                 seed=1)
saveRDS(env_diffT,"fit/env_diffT.rds")

loo(env_sameT,env_diffT)
```

The model with the same thresholds is better

## M0: socio-demographic

```{r}
env_m0 <- brm(data=d_env,
              family=cumulative(),
              prior=prior_m0,
              answer ~ (1|house) + (1|ID) + item +
                gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation),
              iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
              backend="cmdstanr", silent=1,
              seed=1)
saveRDS(env_m0,"fit/env_m0.rds")
```

## M1: predictors

```{r}
prior_m1 <- c(prior(normal(-1.79,2), class=Intercept, coef=1),
              prior(normal(-0.92,2), class=Intercept, coef=2),
              prior(normal(-0.29,2), class=Intercept, coef=3),
              prior(normal(0.29,2), class=Intercept, coef=4),
              prior(normal(0.92,2), class=Intercept, coef=5),
              prior(normal(1.79,2), class=Intercept, coef=6),
              prior(exponential(1), class=sd),
              #beta coefficients = no expectations since sd variables
              prior(normal(0,0.5), class=b, coef=gender_m_w),
              prior(normal(0,0.5), class=b, coef=age_s),
              prior(normal(0,0.5), class=b, coef=income_s),
              prior(normal(0,0.5), class=b, coef=mopol_orientation),
              prior(normal(0,0.5), class=b, coef=moeducation),
              #priors for the predictors, calculated based on the results of the meta-analysis
              prior(normal(0.37,0.2), class=b, coef=si_m_e_s),
              prior(normal(0.33,0.19), class=b, coef=mor_e_s),
              prior(normal(0.35,0.18), class=b, coef=ce_e_s),
              prior(normal(0.35,0.2), class=b, coef=an_e_s))

env_m1 <- brm(data=d_env,
              family=cumulative(),
              prior=prior_m1,
              answer ~ (1|house) + (1|ID) + item +
                gender_m_w + age_s + income_s + mo(education) + mo(pol_orientation) +
                si_m_e_s + mor_e_s + ce_e_s + an_e_s,
              iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
              backend="cmdstanr", silent=1,
              seed=1)
saveRDS(env_m1,"fit/env_m1.rds")
```

## M2: complete model

```{r}
d_env_ce_an <- d %>% 
  pivot_longer(cols=c("ce_env_1","ce_env_2")) %>% 
  rename(answer_ce=value,
         item_ce=name) %>% 
  pivot_longer(cols=starts_with("anger_env")) %>% 
  rename(answer_an=value,
         item_an=name)

env_m2 <- brm(data=d_env_ce_an,
              family=cumulative(),
              prior=prior_m2,
              CE + AN,
              iter=4000, warmup=1000, chains=4, cores=4, threads=threading(4),
              backend = "cmdstanr", silent=1,
              seed=1)

saveRDS(env_m2, "fit/env_m2.rds")
```
